Classical Dynamics of Harmonically Trapped Interacting Particles

### 2017-12-21 draft_10.tex
version submitted to arXiv

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbhp]
\centering
\includegraphics[scale=0.65]{N_5_GS_pre_2_rev-crop.pdf}
\caption{The ground state of a system with $N=5$ particles, shown via the positions of each
  particle, labeled 1 through 5.  At large interactions, the particles position themselves just
  outside the range of interactions of the neighboring particles.}
\label{fig:GS1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=0.9\textwidth]{stationary_and_rotating_frame_t_1-5_5T.pdf}
%
\caption{Upper (lower) panels are snapshots of the cloud configuration in phase space in the
  stationary (rotating) frame, at instants within the first trap period.  Here $N=50$, $F_0=100$,
  and $E=50000$.  }
\label{fig:Breathingfrequency2_0}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=0.9\textwidth]{rotating_frame_t_0-80T.pdf}  
\caption{Phase space snapshots of the cloud dynamics viewed in the rotating frame, at longer
  timescales.  The broadening and rotation observed here are interaction effects: in the
  non-interacting gas the particles are stationary in this frame.}
\label{fig:Breathingfrequency2_1}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbp]
\center
\includegraphics[scale=0.32]{freq_scanF_differentE_log_2_with_oscillation_demo.png}
\caption{Breathing mode frequency measured at different $E$ and $F_0$, for $N=5$ particles. Inset: a
  demonstration of the oscillation of $E_{trap}$, the total potential energy of particles in the
  trap, Since $E_{trap}\propto \sum_{is}{x_i^2}$, it manifests the oscillation behavior of the
  radius $R(t)$ as defined in Eq.~\ref{eq:def_of_R}.  The breathing frequency is obtained from the
  Fourier transform of $R(t)$.}
\label{fig:Breathingfrequency1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbph]
\includegraphics[width=0.95\textwidth]{phase-space_collisioncartoons_01.pdf}
\caption{Schematic diagrams of two-particle collision process in the $X_{\rm rot}$-$P_{\rm rot}$
  plane. 
%
The black filled dots shown locations of the two particles in $X_{\rm rot}$-$P_{\rm rot}$ space
before the collision; the empty dots show locations after the collision.  The dashed lines show the
trajectories during the collision process.  Collistions for different $F_0$ values are shown with
different color.
%
\textbf{(a)} When $F_0$ is large, particles cannot pass each other: during the collision the
momentum changes sign.  Being subject to a constant repulsive force, their trajectories during
collision are curved arcs.  The weaker $F_0$ is, the longer the collision lasts, and hence the more
curved is the corresponding arc.
%
For very strong interactions (red dashed line), particles exchange momenta instantaneously; hence
the red dashed line is straight.  The locations of the particles after this collision are on the
gray line; the angle $\alpha$ made by this gray line is estimated in the text.  The yellow, green,
and blue dashed curves are for successively weaker interactions.  \textbf{(b)} When $F_0$ is small,
particles pass each other, so that the force they feel is discontinuous (changes sign) at the moment
when they cross each other.  Their trajectory in $X_{\rm rot}$-$P_{\rm rot}$ space has a
corresponding turning point.  The dashed bright blue trajectory is identical with the one in (a),
which is the critical situation between passing and bouncing. In that case, particles have zero
relative velocity when they collide. }
\label{fig:Breathingfrequency3}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.45\textwidth]{N_20_freq_vs_E_a.pdf}
\caption{Dependence of breathing frequency, calculated for $N=20$ particles, on energy.  Different
  colors represent different $F_0$ --- red for $F_0=2\times10^3$, yellow for $F_0=3\times10^3$,
  green for $F_0=1\times10^4$, and blue for $F_0=1\times10^5$.  Runs with three different initial
  states are shown in each case.  In each run, the frequency is measured both in the time window
  0-100 (labeled as crosses) and in the time window 900-1000 (labeled as circles). The black dashed
  line is the prediction of Eq.~\eqref{eq:breathingfrequency1}.}
\label{fig:Breathingfrequency4}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.9\textwidth]{two_particles_a_01.pdf}
\caption{Two-particle dynamics.  Left: Effective potential for relative motion.  Center and right:
  evolution of the individual energies of the two particles.  The two panels correspond to internal
  energy $E_i=2$ and $E_i=5$; the interaction is $F_0=1$.  The beat frequency is lower for larger
  $E_i$. }
\label{fig:thermalization2}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{N_5_F0_unknown_E_100_Single_Particle_Energy.pdf}
\caption{Time evolution of single-particle energies in a system of five particles, showing a case of
  inefficient energy redistribution.  The energies of the two particles with highest energy (2 \& 3)
  show a clear beating similar to the two-particle case, as does the pair 4 \& 5, although less
  visible.  The `unpaired' particle 1 performs slow beating dynamics with the center of mass of the
  4 \& 5 pair --- the energy of the particle 1 and of the center of mass of the 4 \& 5 pair undergo
  long-period out-of-phase oscillations, of which approximately the first one-third of one period is
  visible in the time window shown.}
\label{fig:thermalization4}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
%
\includegraphics[width=0.85\textwidth]{Boltzmann_2panels_01-crop.pdf}
%
\caption{The distribution of single-particle energies.  Snapshots of the single-particle energies
  are taken at unit time steps during the simulation, up to $\sim 10000$ time units.  Green curves
  obey the relaxation condition \eqref{eq:thermalizatiton_condition}, and yellow/orange corresponds
  to smaller $F_0$.  The two panels correspond to different particle numbers $N$, but we choose
  total energies such that $E/N$ is identical ($=1000$) in all cases.  Thus the threshold
  \eqref{eq:thermalizatiton_condition} corresponds to the interaction $F_0\sim1000$. }
\label{fig:thermalization5}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\includegraphics[width=0.8\textwidth]{Boltzmann_log_2panels_01-crop.pdf}
%
\caption{\label{fig:thermalization9} Complaring the single-particle energy distribution
  $\rho(\epsilon)$ with the Boltzmann distribution.  The energy values all satisfy the relaxation
  condition \eqref{eq:thermalizatiton_condition}.  The Boltzmann distribution $\propto
  e^{-\beta\epsilon}$, with the temperature defined as $\beta^{-1}= E/N$, is the thick dashed line.}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.55\textwidth]{LLEdistribution_1_11_pre_hot2_screenshot.png}
\caption{The distribution of largest Lyapunov exponents, for $N=5$, $F_0=1000$, at different
  energies.  The color indicates the observed probability of a measured LLE value being found in a
  bin; the bins are of size 0.01.  The average of the distribution decreases monotonically with
  energy.}  \label{fig:LLEdistribution1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tb]
\includegraphics[width=\textwidth]{combine_shapedynamics_a_01-crop.pdf}
\caption{$N=20, F_0=1000$. Time evolution of the shape parameter, $S(t)$, for $N=20$ particles,
  $F_0=1000$.  The three columns correspond to $E<F_0N$ (low-energy or thermalizing regime),
  $E=F_0N$ (intermediate regime), and $E>F_0N$ (high-energy or non-thermalizing regime).  In the top
  row, coarse-grained values (averaged over 10 periods) are shown, for several initial states with
  the same energy in each case.  The lower row shows original data (not averaged) in a window after
  several thousand time steps.  }
\label{fig:time_evolution_of_S}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

>>>>

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbhp]
\centering
\includegraphics[scale=0.65]{ZhiyuPictures/N_5_GS_pre_2_rev-crop.pdf}
\caption{The ground state of a system with $N=5$ particles, shown via the positions of each
  particle, labeled 1 through 5.  At large interactions, the particles position themselves just
  outside the range of interactions of the neighboring particles.}
\label{fig:GS1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=0.9\textwidth]{ZhiyuPictures/stationary_and_rotating_frame_t_1-5_5T.pdf}
%
\caption{Upper (lower) panels are snapshots of the cloud configuration in phase space in the
  stationary (rotating) frame, at instants within the first trap period.  Here $N=50$, $F_0=100$,
  and $E=50000$.  }
\label{fig:Breathingfrequency2_0}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=0.9\textwidth]{ZhiyuPictures/rotating_frame_t_0-80T.pdf}  
\caption{Phase space snapshots of the cloud dynamics viewed in the rotating frame, at longer
  timescales.  The broadening and rotation observed here are interaction effects: in the
  non-interacting gas the particles are stationary in this frame.}
\label{fig:Breathingfrequency2_1}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbp]
\center
\includegraphics[scale=0.32]{ZhiyuPictures/freq_scanF_differentE_log_2_with_oscillation_demo.png}
\caption{Breathing mode frequency measured at different $E$ and $F_0$, for $N=5$ particles. Inset: a
  demonstration of the oscillation of $E_{trap}$, the total potential energy of particles in the
  trap, Since $E_{trap}\propto \sum_{is}{x_i^2}$, it manifests the oscillation behavior of the
  radius $R(t)$ as defined in Eq.~\ref{eq:def_of_R}.  The breathing frequency is obtained from the
  Fourier transform of $R(t)$.}
\label{fig:Breathingfrequency1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbph]
\includegraphics[width=0.95\textwidth]{ZhiyuPictures/phase-space_collisioncartoons_01.pdf}
\caption{Schematic diagrams of two-particle collision process in the $X_{\rm rot}$-$P_{\rm rot}$
  plane. 
%
The black filled dots shown locations of the two particles in $X_{\rm rot}$-$P_{\rm rot}$ space
before the collision; the empty dots show locations after the collision.  The dashed lines show the
trajectories during the collision process.  Collistions for different $F_0$ values are shown with
different color.
%
\textbf{(a)} When $F_0$ is large, particles cannot pass each other: during the collision the
momentum changes sign.  Being subject to a constant repulsive force, their trajectories during
collision are curved arcs.  The weaker $F_0$ is, the longer the collision lasts, and hence the more
curved is the corresponding arc.
%
For very strong interactions (red dashed line), particles exchange momenta instantaneously; hence
the red dashed line is straight.  The locations of the particles after this collision are on the
gray line; the angle $\alpha$ made by this gray line is estimated in the text.  The yellow, green,
and blue dashed curves are for successively weaker interactions.  \textbf{(b)} When $F_0$ is small,
particles pass each other, so that the force they feel is discontinuous (changes sign) at the moment
when they cross each other.  Their trajectory in $X_{\rm rot}$-$P_{\rm rot}$ space has a
corresponding turning point.  The dashed bright blue trajectory is identical with the one in (a),
which is the critical situation between passing and bouncing. In that case, particles have zero
relative velocity when they collide. }
\label{fig:Breathingfrequency3}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.45\textwidth]{ZhiyuPictures/N_20_freq_vs_E_a.pdf}
\caption{Dependence of breathing frequency, calculated for $N=20$ particles, on energy.  Different
  colors represent different $F_0$ --- red for $F_0=2\times10^3$, yellow for $F_0=3\times10^3$,
  green for $F_0=1\times10^4$, and blue for $F_0=1\times10^5$.  Runs with three different initial
  states are shown in each case.  In each run, the frequency is measured both in the time window
  0-100 (labeled as crosses) and in the time window 900-1000 (labeled as circles). The black dashed
  line is the prediction of Eq.~\eqref{eq:breathingfrequency1}.}
\label{fig:Breathingfrequency4}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.9\textwidth]{ZhiyuPictures/two_particles_a_01.pdf}
\caption{Two-particle dynamics.  Left: Effective potential for relative motion.  Center and right:
  evolution of the individual energies of the two particles.  The two panels correspond to internal
  energy $E_i=2$ and $E_i=5$; the interaction is $F_0=1$.  The beat frequency is lower for larger
  $E_i$. }
\label{fig:thermalization2}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{ZhiyuPictures/N_5_F0_unknown_E_100_Single_Particle_Energy.pdf}
\caption{Time evolution of single-particle energies in a system of five particles, showing a case of
  inefficient energy redistribution.  The energies of the two particles with highest energy (2 \& 3)
  show a clear beating similar to the two-particle case, as does the pair 4 \& 5, although less
  visible.  The `unpaired' particle 1 performs slow beating dynamics with the center of mass of the
  4 \& 5 pair --- the energy of the particle 1 and of the center of mass of the 4 \& 5 pair undergo
  long-period out-of-phase oscillations, of which approximately the first one-third of one period is
  visible in the time window shown.}
\label{fig:thermalization4}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
%\includegraphics[scale=0.2]{ZhiyuPictures/distribution1_scanF0_E=100_3.eps}
%\caption{Distribution scan $F_0$, $E=100$, $N=5$, $\sigma=1$ {\color{red}{Run again!}} }
% \includegraphics[width=0.32\textwidth]{ZhiyuPictures/N=5_energydistribution_500_400_Font16.eps} 
% \includegraphics[width=0.32\textwidth]{ZhiyuPictures/N=10_energydistribution_500_400_Font16.eps}
% \includegraphics[width=0.32\textwidth]{ZhiyuPictures/N=20_energydistribution_500_400_Font16.eps}
%
\includegraphics[width=0.85\textwidth]{ZhiyuPictures/Boltzmann_2panels_01-crop.pdf}
%
\caption{The distribution of single-particle energies.  Snapshots of the single-particle energies
  are taken at unit time steps during the simulation, up to $\sim 10000$ time units.  Green curves
  obey the relaxation condition \eqref{eq:thermalizatiton_condition}, and yellow/orange corresponds
  to smaller $F_0$.  The two panels correspond to different particle numbers $N$, but we choose
  total energies such that $E/N$ is identical ($=1000$) in all cases.  Thus the threshold
  \eqref{eq:thermalizatiton_condition} corresponds to the interaction $F_0\sim1000$. }
\label{fig:thermalization5}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\includegraphics[width=0.8\textwidth]{ZhiyuPictures/Boltzmann_log_2panels_01-crop.pdf}
%
\caption{\label{fig:thermalization9} Complaring the single-particle energy distribution
  $\rho(\epsilon)$ with the Boltzmann distribution.  The energy values all satisfy the relaxation
  condition \eqref{eq:thermalizatiton_condition}.  The Boltzmann distribution $\propto
  e^{-\beta\epsilon}$, with the temperature defined as $\beta^{-1}= E/N$, is the thick dashed line.}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tb]
\centering
%\includegraphics[scale=0.4]{ZhiyuPictures/LLEdistribution_1_11_pre_hot1_screenshot.png}
\includegraphics[width=0.55\textwidth]{ZhiyuPictures/LLEdistribution_1_11_pre_hot2_screenshot.png}
\caption{The distribution of largest Lyapunov exponents, for $N=5$, $F_0=1000$, at different
  energies.  The color indicates the observed probability of a measured LLE value being found in a
  bin; the bins are of size 0.01.  The average of the distribution decreases monotonically with
  energy.}  \label{fig:LLEdistribution1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

@@

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tb]
\includegraphics[width=\textwidth]{ZhiyuPictures/combine_shapedynamics_a_01-crop.pdf}
\caption{$N=20, F_0=1000$. Time evolution of the shape parameter, $S(t)$, for $N=20$ particles,
  $F_0=1000$.  The three columns correspond to $E<F_0N$ (low-energy or thermalizing regime),
  $E=F_0N$ (intermediate regime), and $E>F_0N$ (high-energy or non-thermalizing regime).  In the top
  row, coarse-grained values (averaged over 10 periods) are shown, for several initial states with
  the same energy in each case.  The lower row shows original data (not averaged) in a window after
  several thousand time steps.  }
\label{fig:time_evolution_of_S}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%
